---
format: typst
---

### Easy

#### E1

$y_i \sim Normal(\mu, \sigma)$ is the likelihood.

#### E2

There are two parameters: $\mu$ and $\sigma$.

#### E3

$$
\begin{align*}
[\mu,\sigma|y] \propto &\prod_i^nNormal(y_i|\mu,\sigma)\times\\
&Normal(\mu|0,10)\times\\
&Uniform(\sigma|0,10)
\end{align*}
$$

#### E4

$\mu_i = \alpha + \beta x_i$ is the linear model.

#### E5

There are 3 parameters: $\alpha$, $\beta$, and $\sigma$.

### Medium

#### M1

```{r}
sigma <- runif(1000, 0, 10)
mu <- rnorm(1000, 0, 10)
y <- rnorm(mu, sigma)
hist(sigma)
hist(mu)
hist(y)
```

#### M2

```{r}
alist(
  y ~ dnorm(mu, sigma),
  mu ~ dnorm(0, 10),
  sigma ~ dunif(0,10)
)
```

#### M3

$$
\begin{align*}
y_i &\sim Normal(\mu_i, \sigma)\\
\mu_i &= \alpha + \beta x_i\\
\alpha &\sim Normal(0, 50)\\
\beta &\sim Uniform(0, 10)\\
\sigma &\sim Uniform(0, 50)
\end{align*}
$$

#### M4 

I will standardize height across all three years to make choosing priors easier. I will also code the first year as -1, the second as 0 and the third year as 1. This will make $\alpha$ very close to zero. Setting $\beta$ to likely be positive (kids tend to grow). And $\log(\sigma)$ has a distribution around 1. These give a prior predictive distribution that seems reasonable

```{r}
# center year so that it is (-1, 0, 1)
# standardize height measurements so that they have a mean of 0 and a stdv of 1
n <- 100
a <- rnorm(n,0, .5)
b <- rnorm(n,.5, .5)
y <- matrix(nrow = n, ncol = 3)
sigma <- rnorm(n, 0, .25)
for(i in 1:3){
  y[,i] <- rnorm(n, a + b*(i-2), exp(sigma))
}

hist(y)
abline(v = quantile(y,c(.025, .975)),
       col = "red")

year <- c(-1,0,1)
plot(apply(y,2,mean) ~ year, type = "l")
```



$$
\begin{align*}
height_{i} &\sim Normal(\mu_{i}, \sigma) \\
\mu_i &= \alpha + \beta\times year_i \\
\alpha &\sim Normal(0, .5) \\
\beta &\sim Normal(.5,.5) \\
\log(\sigma) &\sim Normal(0, .25) \\
\end{align*}
$$

#### M5

It wouldn't with how I decided to standardize the variables, but in the spirit of the question, I would amend $\alpha$ to have a mean of 120 with an sd of 10. And let's say they grow an average of 2 ish centimeters per year. Giving $\log(\sigma)$ a mean 0f 0 and sd of 1 seems to give reasonable results from the prior predictive. 

```{r}

a <- rnorm(n,120, 10)
b <- rnorm(n,2,.5)
sigma <- rnorm(n,0, 1)

y <- matrix(nrow = n, ncol = 3)
for(i in 1:3){
  y[,i] <- rnorm(n, a + b*(i-2), exp(sigma))
}

hist(y)
plot(apply(y, 2, mean) ~ year, type = "l")

```
$$
\begin{align*}
height_{i} &\sim Normal(\mu_{i}, \sigma) \\
\mu_i &= \alpha + \beta\times year_i \\
\alpha &\sim Normal(120, 10) \\
\beta &\sim Normal(2,.5) \\
\log(\sigma) &\sim Normal(0, 1) \\
\end{align*}
$$

#### M6

A variance of 64 is a standard deviation of 8, and my prior is predicting more variation than that, so I can probably tighten it down. Going to $\log(\sigma) \sim Normal(0, .5)$ seems to work well.
```{r}

hist(replicate(1e4, max(exp(rnorm(100, 0, .5)))))
```
$$
\begin{align*}
height_{i} &\sim Normal(\mu_{i}, \sigma) \\
\mu_i &= \alpha + \beta\times year_i \\
\alpha &\sim Normal(120, 10) \\
\beta &\sim Normal(2,.5) \\
\log(\sigma) &\sim Normal(0, .5) \\
\end{align*}
$$

